---
title: "INFO510_2020"
author: "Malay"
date: "5/26/2020"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

# 1st day

## Vectors
```{r}
cat("Hello world")


x = c(1,2,3,4, "hello")
x <- c(1,2,3,4) 
x <- c(1:100)
x + 2

x[1]
x[100]
x
class(x)

y <- c("a","b","c",1)
y
class(y)
```

## Lists

```{r}
l <- list()
l[[1]] <- 1
l[[2]] <- "Hello"

v <- vector()
v[1] <- 1
v[2] <- "Hello"

class(l)
class(v)

str(l)

```

## Data frame

```{r}
data(iris)
View(iris)
str(iris)

dim(iris)
```


### Subsetting data frame

```{r}
colnames(iris)
row.names(iris)

# Particular column
iris$Species
iris[ ,5 ]

# Particular cell
iris[1,2]

# Particular row
iris[10, ]

l[[3]] <- iris
str(l)

l <- list()
l[[1]] <- iris$Sepal.Length
l[[2]] <- iris$Sepal.Width
l[[3]] <- iris
str(l)

l[[1]][2]
l[[2]][2]

# getting the species column from the 3rd element of the list
l[[3]][, 5]
```

## Factor

```{r}
unique(iris$Species)
class(iris$Species)
levels(iris$Species)

myiris <- split(iris, iris$Species)

x <- factor(c("yes", "yes", "no"), levels = c("yes","no"))
x
```

## Looping

```{r}
for (i in 1:length(myiris) ) {
  df <- myiris[[i]]
  print (mean( df$Sepal.Length))
}
mean()
```

### Apply family

```{r}

# lapply
results <- lapply( myiris, function(x) {
  mean(x$Sepal.Length)
} )  

results[[1]]
results[["setosa"]]

# apply
apply_result <- apply(iris[ , c(1:4)], MARGIN = 2, mean)
apply_result
```

## Logical data type

```{r}
v <- c(1:100)
v <- v+2
v [v>50 & v<90]
v <- v [ v != 59]
v

# subsetting iris for only numeric columns
str(iris)

only_num <- vector()

for (i in 1:dim(iris)[2])  {
  v <- iris[, i]
  only_num <- c(only_num, class(v) == "numeric")
}
only_num
only_num_iris <- iris[, only_num]
View(only_num_iris)
apply(only_num_iris, MARGIN = 2, mean)

#----------------


```

## Converting one type to another

```{r}
v <- c(1:2, "a")
v <- c(1:2)
class(v)
vc <- as.character(v)
class(vc)
vc
f <- iris$Species
class(f)
as.numeric(f)
as.character(f)
iris[iris$Species == "setosa", ]
```

## Sorting

```{r}
o <- order(iris$Petal.Length)
ordered_vector <- iris$Petal.Length[o] 
ordered_vector

# Ordering data frame
ordered_iris <- iris[o, ]
View(ordered_iris)
```

## Matrix

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
dim(m)
m
dim(m) <- c(3,2)
m
```

## Traspose
```{r}
trasposed_iris <- as.data.frame(t(iris))

```

## NA

```{r}
x <- c(1,2, NA, 10, 3)
mean(x)
!is.na(x)
x <- x[!is.na(x)]
x
```

## Read and write data in R

```{r}
View(iris)

write.table(iris, file="myiris.txt", row.names = F, 
            quote = F, col.names = T, sep="\t")

myiris <- read.table("myiris.txt", header = T)


```

# Plotting

```{r}
plot(iris)
x <- iris$Sepal.Length
y <- iris$Sepal.Width

{
#png(file = "testplot.png")
par(mfrow=c(1,2))
plot(x,y, main = "Iris data", pch=16)
  points(iris$Petal.Length,iris$Petal.Width, col="red", pch=16)
legend("topleft", legend=c("Sepal","Petal"), pch=16, col=c("black","red") )


plot(x,y, main = "Iris data", pch=16)
  points(iris$Petal.Length,iris$Petal.Width, col="red", pch=16)
legend("topleft", legend=c("Sepal","Petal"), pch=16, col=c("black","red") )
#dev.off()
par(mfrow=c(1,1))
}

plot(x,y, main = "Iris data", pch=16, col=iris$Species)

# ggplot
library(ggplot2)
library(cowplot)
library(patchwork)
p <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, colour = Species)) + geom_point() + theme_set(theme_cowplot())
print(p)

# Multiplot in ggplot
p1 <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, colour = Species)) + geom_point() + theme_set(theme_cowplot()) +ggtitle("Plot 1\ntitle")
p2 <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, colour = Species)) + geom_point() + theme_set(theme_cowplot()) 

combined_plot <- plot_grid(p1, p2, nrow=1) # cowplot
combined_plot

p1 <- p1 + theme_dark()


patched_p <- p1 + p2 # patchwork
patched_p

x <- 2
x
print(x)
```


## Histogram

```{r}
# Base graphics
head(iris$Petal.Length)
hist( iris$Petal.Length, breaks = 50)
plot( density(iris$Petal.Length))

{plot (density (rnorm(500)))
abline(v=1)
abline(v=-1)
}
rn <- rnorm(500)
mean(rn)
sd(rn)

## ggplot
g <- ggplot(iris, aes(x=Petal.Width))
g <- g + geom_histogram(bins = 50)
g
```


## Boxplot

```{r}
data("airquality")
View(airquality)

boxplot(airquality)

n <- rnorm(101)
n <- n[order(n)]
median(n)
n
quantile(n, 0.25)
quantile(n, 0.75)

summary(airquality)
table(iris$Species)

##ggplot

p <- ggplot(airquality, aes(x=) )

```

# Satistics

## Correlation
```{r}
plot (airquality$Solar.R, airquality$Temp)
s <- cor.test(airquality$Solar.R, airquality$Temp, method = )
s$p.value
s$estimate

```

# Regression

```{r}
{plot(airquality$Solar.R, airquality$Temp)
m <- lm(airquality$Temp ~ airquality$Solar.R)
summary(m)
abline(m)}
```

# Kable

```{r}
knitr::kable(airquality)
```



# Regular expression

1. "\n" -newline
2. "\t" -tab
3. "\d" - number
4. "\S" - Nonspace
5. "\s" - space
6. "+" - One or more
7. "*" - None or more
8. "." - Any character
9. "^" - beginning
10 "$" - end

## Regular expression in R

```{r}
iris$Species
iris$Species == "setosa"
iris$Species[ grep("^v", iris$Species)]
grepl("^v", iris$Species)

```













